
\chapter*{Sustainable Development Goals}
%\addcontentsline{toc}{chapter}{Abstract}
This dissertation takes into account the Sustainable Development Goals defined by the United Nations.
It directly addresses 4 out of 17 goals.
Those are:
\begin{enumerate}
    \item 8: Decent Work and Economic Growth
    \item 9: Industry, Innovation, and Infrastructure
    \item 11: Sustainable Cities and Communities
    \item 12: Responsible Consumption and Production
\end{enumerate}
Table~\ref{tab:ODS_table} summarizes the targets this work contributes to and its indicators.


\begin{table}[H]
    \begin{tabular}{|c|c|c|c|}
        \hline
        \textbf{SDG} & \textbf{Target} & \textbf{Contribution} & \textbf{Indicators} \\ \hline
        8            & 8.2         &                       &                     \\ \hline
        9            &               &                       &                     \\ \hline
        11            &               &                       &                     \\ \hline
        12&               &                       &                     \\ \hline
    \end{tabular}\label{tab:ODS_table}
\end{table}


Specific Targets Within Identified SDGs

SDG 8: Decent Work and Economic Growth
- **Target 8.2:** Achieve higher levels of economic productivity through diversification, technological upgrading, and innovation.
- *Contribution:* Your project enhances economic productivity by leveraging computer vision to improve the efficiency and capabilities of 5G networks, fostering technological innovation and creating high-tech job opportunities.

SDG 9: Industry, Innovation, and Infrastructure
- **Target 9.1:** Develop quality, reliable, sustainable, and resilient infrastructure.
- *Contribution:* Integrating computer vision with 5G networks can significantly enhance infrastructure reliability and performance, supporting sustainable development.
- **Target 9.4:** Upgrade infrastructure and retrofit industries to make them sustainable, with increased resource-use efficiency and greater adoption of clean and environmentally sound technologies.
- *Contribution:* By optimizing network performance and reducing energy consumption through advanced computer vision techniques, your project promotes resource efficiency and the use of environmentally friendly technology.

SDG 11: Sustainable Cities and Communities
- **Target 11.6:** Reduce the adverse per capita environmental impact of cities.
- *Contribution:* Enhanced 5G networks can support smarter city management and reduce environmental impact by enabling efficient traffic management, pollution monitoring, and resource utilization through real-time data processing.

SDG 12: Responsible Consumption and Production
- **Target 12.2:** Achieve the sustainable management and efficient use of natural resources.
- *Contribution:* Computer vision applications can optimize the supply chain and production processes, reducing waste and promoting the efficient use of resources.

Contribution to Improving Quality of Life

- **Economic Growth:** By driving innovation and technological advancements, your project supports job creation and economic development.
- **Environmental Protection:** Optimized 5G networks can lead to more efficient use of resources and lower emissions, contributing to environmental sustainability.
- **Social Equality:** Improved connectivity and infrastructure can provide equitable access to technology and services, reducing the digital divide.
- **Urban Development:** Smart city initiatives supported by enhanced 5G networks can improve urban living conditions through better management of services and resources.

Performance Indicators

To measure the impact of your project, you can establish the following performance indicators:
- **Economic Indicators:** Job creation statistics in the tech sector, GDP growth related to tech industries, and levels of investment in 5G and computer vision technologies.
- **Infrastructure Efficiency:** Network reliability metrics, energy consumption rates of 5G networks, and reduction in downtime or service interruptions.
- **Environmental Impact:** Reduction in emissions due to more efficient network operations, levels of resource use efficiency, and waste reduction in production processes.
- **Urban Metrics:** Improvements in traffic management efficiency, reduction in urban pollution levels, and enhanced service delivery in smart city applications.